---
title: "Benchmark report"
author: "Akos Hajdu"
date: '2017'
output:
  html_document:
    code_folding: hide
    number_sections: yes
    toc: no
    toc_depth: 2
  pdf_document:
    highlight: zenburn
    number_sections: yes
    toc: no
    toc_depth: 2
classoption: a4paper
---

# Initialization

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stringr)
library(forcats)
```

```{r}
# Common variables
csv_path = "../log_20170310_101235.csv" # Path of the data
timeout_ms = 10 * 60 * 1000 # Timeout used during the measurements

# Themes
theme_rotate_x <- theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
theme_noticks <- theme(axis.ticks = element_blank())
```

## Load data

```{r}
d <- read_csv(
  csv_path,
  col_types = cols(
    Model = col_character(),
    Vars = col_integer(),
    Size = col_integer(),
    Domain = col_character(),
    Refinement = col_character(),
    InitPrec = col_character(),
    Search = col_character(),
    PredSplit = col_character(),
    Safe = col_character(),
    TimeMs = col_integer(),
    Iterations = col_integer(),
    ArgSize = col_integer(),
    ArgDepth = col_integer(),
    CexLen = col_integer()
  )
)
```

## Clean data

```{r}
# Create factors
d$Domain <- factor(d$Domain)
d$Refinement <- factor(d$Refinement)
d$InitPrec <- factor(d$InitPrec)
d$Search <- factor(d$Search)
d$PredSplit <- factor(d$PredSplit)

# Convert 'Safe' to logical (and remove exceptions).
ex_rows <- str_detect(d$Safe, "\\[EX\\]")
exs <- sum(ex_rows, na.rm = T)
if (exs > 0) warning(paste(c(exs, " rows contain exceptions that are converted to NAs.")))
d$Safe <- as.logical(d$Safe)

# Create 'Config' column.
d <- d %>% mutate(Config = paste(
  substr(d$Domain, 1, 1),
  substr(d$Refinement, 1, 1),
  substr(d$InitPrec, 1, 1),
  substr(d$Search, 1, 1),
  ifelse(is.na(d$PredSplit), "", substr(d$PredSplit, 1, 1)),
  sep = ""
))

# Trim name, separate 'Model' into 'Name' and 'Type'.
d$Model <- as.factor(gsub("models/", "", d$Model))
d <- d %>% separate(Model, into = c("Type", "Name"), sep = "/", remove = F)
```

# Analysis

## Quick summary


## Consistency of results
```{r}
d_model_config <- d %>% group_by(Config, Model) %>%
  summarise(
    ResultCount = sum(!is.na(Safe)),
    SafeCount = sum(Safe, na.rm = T),
    Consistent = SafeCount == 0 || SafeCount == ResultCount,
    TimeAvg = mean(TimeMs),
    TimeRSD = sd(TimeMs) / mean(TimeMs))
d_model <- d_model_config %>% group_by(Model) %>%
  summarise(
    ResultCount = sum(ResultCount),
    SafeCount = sum(SafeCount),
    Consistent = SafeCount == 0 || SafeCount == ResultCount)
```

There are **`r sum(!d_model_config$Consistent)`** cases where different executions of a configuration yielded different results for a model.
```{r}
d_model_config %>% filter(Consistent == F) %>% select(Config, Model)
```

There are **`r sum(!d_model$Consistent)`** cases where different executions of a configuration or different configurations yielded different results for a model.
```{r}
d_model %>% filter(Consistent == F) %>% select(Model)
```

## Overview

### Relative Standard Deviation

```{r}
ggplot(d_model_config) +
  geom_histogram(aes(TimeRSD)) +
  labs(title = "Overall distribution of the RSD")
```

The maximum RSD is **`r max(d_model_config$TimeRSD, na.rm = T)`**, and 95% of the measurements has a lower RSD than `r quantile(d_model_config$TimeRSD, .95, na.rm = T)`.

```{r}
ggplot(d_model_config, aes(Model, Config)) +
  geom_tile(aes(fill = TimeRSD), color = "black") +
  scale_fill_gradient(low = "green", high = "red", na.value = "white") +
  theme_noticks + theme_rotate_x +
  labs(title = "Individual RSD for each configuration and model")

```

### Average execution time
```{r}
ggplot(d_model_config, aes(Model, Config)) +
  geom_tile(aes(fill = log(TimeAvg)), color = "black") +
  scale_fill_gradient(low = "green", high = "red", na.value = "white") +
  theme_noticks + theme_rotate_x
```

