---
title: "Benchmark report"
params:
  csv_path: "log_20170412_170912.csv"
  timeout_ms: 600000
output:
  html_document:
    code_folding: hide
    number_sections: yes
    toc: no
    toc_depth: 2
  pdf_document:
    highlight: zenburn
    number_sections: yes
    toc: no
    toc_depth: 2
classoption: a4paper
---

This is a benchmark report for the file `r params$csv_path`.

# Initialization

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stringr)
library(forcats)
library(lubridate)
library(corrgram)
```

```{r}
# Themes
theme_rotate_x <- theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
theme_noticks <- theme(axis.ticks = element_blank())

# Function for pairwise plots
pairwise_plot <- function(data, compare_col, group_col, val1, val2, other_cols, color_by,
                          na_fill = NA, log_scale = F, bin_width = 1, pt_position = "identity") {
  all_cols <- c(group_col, compare_col, other_cols)
  data_selected <- data %>% select(one_of(all_cols))
  data_selected[[compare_col]][is.na(data_selected[[compare_col]])] <- na_fill
  data_joined <- inner_join(
    data_selected %>% filter_(lazyeval::interp(quote(x == y), x=as.name(group_col), y=val1)),
    data_selected %>% filter_(lazyeval::interp(quote(x == y), x=as.name(group_col), y=val2)),
    by = other_cols)
  plot <- ggplot(data_joined, aes_string(paste(compare_col, ".x", sep = ""), paste(compare_col, ".y", sep = "")))
  if (log_scale) {plot <- plot + scale_x_log10() + scale_y_log10()}
  plot <- plot + geom_bin2d(binwidth = bin_width)
  plot <- plot + scale_fill_gradient(low = "gray", high = "black")
  plot <- plot + geom_point(aes_string(color = color_by), position = pt_position, size = 2)
  plot <- plot + geom_abline()
  plot <- plot + labs(title = paste("Comparing", compare_col, "for", group_col), x = val1, y = val2)
  plot
}
```

## Load data

```{r}
d <- read_csv(
  params$csv_path,
  col_types = cols(
    Expected = col_logical(),
    Model = col_character(),
    Prop = col_character(),
    Domain = col_character(),
    InitPrec = col_character(),
    PredSplit = col_character(),
    Refinement = col_character(),
    Search = col_character(),
    Safe = col_character(),
    TimeMs = col_integer(),
    Iterations = col_integer(),
    ArgSize = col_integer(),
    ArgDepth = col_integer(),
    CexLen = col_integer(),
    Vars = col_integer(),
    Size = col_integer()
  )
)
```

## Clean data

```{r}
# Drop unneeded columns
d <- d %>% select(-Expected, -Prop)

# Create factors
d$Domain <- factor(d$Domain)
d$Refinement <- factor(d$Refinement)
d$InitPrec <- factor(d$InitPrec)
d$Search <- factor(d$Search)
d$PredSplit[is.na(d$PredSplit)] <- "_"
d$PredSplit <- factor(d$PredSplit)

# Convert 'Safe' to logical (and remove exceptions).
empty_rows <- sum(is.na(d$Safe))
if (empty_rows > 0) warning(paste(c(empty_rows, " rows contain missing values (no timeout, no exception)")))
ex_rows <- str_detect(d$Safe, "\\[EX\\]")
exs <- sum(ex_rows, na.rm = T)
if (exs > 0) warning(paste(c(exs, " rows contain exceptions that are converted to NAs.")))
d$Safe <- as.logical(d$Safe)

# Create 'Config' column.
d <- d %>% mutate(Config = paste(
  substr(d$Domain, 1, 1),
  substr(d$Refinement, 1, 1),
  substr(d$InitPrec, 1, 1),
  substr(d$Search, 1, 1),
  ifelse(is.na(d$PredSplit), "", substr(d$PredSplit, 1, 1)),
  sep = ""
))

# Trim name, separate 'Model' into 'Name' and 'Type'.
d$Model <- as.factor(gsub("../models/sts/", "", d$Model))
d <- d %>% separate(Model, into = c("Type", "Name"), sep = "/", remove = F)
```

# Overview

## Summary

```{r}
n_meas_total <- nrow(d)
n_meas_succ <- nrow(d %>% filter(!is.na(Safe)))
n_models_total <- length(unique(d$Model))
n_configs_total <- length(unique(d$Config))
total_time <- seconds_to_period(sum(ifelse(is.na(d$TimeMs), params$timeout_ms, d$TimeMs)) / 1000)

d_model_config <- d %>%
  group_by(Config, Model, Domain, Refinement, InitPrec, Search, PredSplit, Type) %>%
  summarise(
    ResultCount = sum(!is.na(Safe)),
    Succ = ResultCount > 0,
    SafeCount = sum(Safe, na.rm = T),
    Consistent = SafeCount == 0 || SafeCount == ResultCount,
    TimeAvg = mean(TimeMs),
    TimeRSD = ifelse(ResultCount == 1, 0, sd(TimeMs) / mean(TimeMs)),
    IterAvg = mean(Iterations))
d_model <- d_model_config %>% group_by(Model, Type) %>%
  summarise(
    ResultCount = sum(ResultCount),
    Succ = ResultCount > 0,
    SafeCount = sum(SafeCount),
    Consistent = SafeCount == 0 || SafeCount == ResultCount,
    TimeMin = min(TimeAvg, na.rm = T),
    ConfigMin = ifelse(is.na(TimeMin), "", Config[which.min(TimeAvg)]))
d_model$ConfigMin[d_model$ConfigMin == ""] <- NA

n_models_succ <- length(unique((d_model %>% filter(Succ))$Model))
n_configs_succ <- length(unique((d_model_config %>% filter(Succ))$Config))
```

- There are **`r n_models_total` models** and **`r n_configs_total` configurations**, giving **`r n_models_total * n_configs_total` measurements**, with **`r nrow(d_model_config %>% filter(Succ))` being successful** with the timeout limit of `r params$timeout_ms/1000` s (`r nrow(d_model_config %>% filter(Succ)) / (n_models_total * n_configs_total)` success rate).
    - **`r n_models_succ` / `r n_models_total` models** were verified by at least one configuration.
    - **`r n_configs_succ` / `r n_configs_total` configurations** verified at least one model.
- With the repeated measurements included, there are a total number of **`r n_meas_total` measurement points** with **`r n_meas_succ` successful** executions.
- Total time of the measurements is **`r total_time$day` days, `r total_time$hour` hours and `r total_time$minute` minutes**, assuming that the missing values timed out.
    - The maximum (non-timeout) execution time is **`r max(d$TimeMs, na.rm = T) / 1000` s**.

## Consistency of results

There are **`r sum(!d_model_config$Consistent)`** cases where different executions of the same configuration yielded different results for the same model.
```{r}
d_model_config %>% filter(!Consistent) %>% ungroup() %>% select(Config, Model)
```

There are **`r sum(!d_model$Consistent)`** cases where different executions of the same configuration or different configurations yielded different results for the same model.
```{r}
d_model %>% filter(!Consistent) %>% select(Model)
```

## Overview of models and configurations

```{r, fig.height = 4}
ggplot(d_model_config) +
  geom_bar(aes(x = Model, fill = Succ)) +
  scale_x_discrete(drop = F) +
  theme_rotate_x +
  labs(title = "Number of configurations that verified a model")
```

```{r, fig.height = 4}
ggplot(d_model_config) +
  geom_bar(aes(x = Config, fill = Succ)) +
  scale_x_discrete(drop = F) +
  theme_rotate_x +
  labs(title = "Number of models verified by a configuration")
```

```{r, fig.height = 4}
ggplot(d_model_config %>% filter(Succ)) +
  geom_bar(aes(x = Config, fill = Type), position = "dodge") +
  scale_x_discrete(drop = F) +
  theme_rotate_x +
  labs(title = "Number of models verified by a configuration in each type")
```

```{r}
ggplot(d_model_config, aes(Config, Model)) +
  geom_tile(aes(fill = ResultCount), color = "black") +
  scale_fill_gradient(low = "red", high = "green", na.value = "white") +
  theme_noticks + theme_rotate_x +
  labs(title = "Number of successful executions")
```

# Details

## Verification success

```{r, fig.height = 2}
ggplot(d_model_config) +
  geom_bar(aes(x = Domain, fill = Succ), position = "fill") +
  facet_wrap(~ Type) +
  coord_flip() +
  labs(title = "Success rate of domains")
```

```{r, fig.height = 2}
ggplot(d_model_config) +
  geom_bar(aes(x = Refinement, fill = Succ), position = "fill") +
  facet_wrap(~ Type) +
  coord_flip() +
  labs(title = "Success rate of refinements")
```

```{r, fig.height = 2}
ggplot(d_model_config) +
  geom_bar(aes(x = InitPrec, fill = Succ), position = "fill") +
  facet_wrap(~ Type) +
  coord_flip() +
  labs(title = "Success rate of initial precisions")
```

```{r, fig.height = 2}
ggplot(d_model_config) +
  geom_bar(aes(x = Search, fill = Succ), position = "fill") +
  facet_wrap(~ Type) +
  coord_flip() +
  labs(title = "Success rate of searches")
```

```{r, fig.height = 2}
ggplot(d_model_config) +
  geom_bar(aes(x = PredSplit, fill = Succ), position = "fill") +
  facet_wrap(~ Type) +
  coord_flip() +
  labs(title = "Success rate of predicate splits")
```

## Execution time

### Relative Standard Deviation

```{r, fig.height = 2}
ggplot(d_model_config) +
  geom_histogram(aes(TimeRSD)) +
  labs(title = "Overall distribution of the RSD")
```

The maximum RSD is **`r max(d_model_config$TimeRSD, na.rm = T)`**, and 95% of the measurements have a lower RSD than `r quantile(d_model_config$TimeRSD, .95, na.rm = T)`.

```{r}
ggplot(d_model_config, aes(Config, Model)) +
  geom_tile(aes(fill = TimeRSD), color = "black") +
  scale_fill_gradient(low = "green", high = "red", na.value = "white") +
  theme_noticks + theme_rotate_x +
  labs(title = "Individual RSD for each configuration and model")
```

### Average execution time

```{r, fig.height = 3}
ggplot(d %>% filter(!is.na(TimeMs))) +
  geom_histogram(aes(TimeMs)) +
  labs(title = "Distribution of execution time")
summary(d$TimeMs)
```

```{r}
ggplot(d %>% filter(!is.na(TimeMs))) +
  geom_boxplot(aes(Type, log10(TimeMs))) +
  coord_flip() +
  labs(title = "Distribution of execution time for each type")
```



```{r}
ggplot(d %>% filter(!is.na(TimeMs))) +
  geom_boxplot(aes(Model, log10(TimeMs))) + coord_flip() +
  labs(title = "Distribution of execution time for each model")
```

```{r}
ggplot(d_model_config, aes(Config, Model)) +
  geom_tile(aes(fill = log10(TimeAvg)), color = "black") +
  scale_fill_gradient(low = "green", high = "red", na.value = "white") +
  theme_noticks + theme_rotate_x +
  labs(title = "Heatmap of execution time for each model/configuration")
```

```{r, fig.height = 4}
ggplot(d_model %>% filter(!is.na(ConfigMin))) +
  geom_bar(aes(x = ConfigMin, fill = Type)) +
  theme_rotate_x +
  labs(title = "Number of models where a configuration was the fastest")
```

### Pairwise comparisons
```{r}
pairwise_plot(d, "TimeMs", "Domain", "PRED", "EXPL",
              c("Type", "Model", "Refinement", "InitPrec", "Search"), "Type",
              log_scale = T, bin_width = .25, na_fill = params$timeout_ms, pt_position = "jitter")
```

## Iterations

```{r}
ggplot(d_model_config, aes(Config, Model)) +
  geom_tile(aes(fill = IterAvg), color = "black") +
  scale_fill_gradient(low = "green", high = "red", na.value = "white") +
  theme_noticks + theme_rotate_x +
  labs(title = "Heatmap of iterations for each model/configuration")
```

```{r}
pairwise_plot(d, "Iterations", "Refinement", "BW_CRAIG_ITP", "FW_CRAIG_ITP",
              c("Type", "Model", "Domain", "InitPrec", "Search", "PredSplit"), "Type",
              log_scale = F, bin_width = 2, pt_position = "jitter")
pairwise_plot(d, "Iterations", "Refinement", "BW_CRAIG_ITP", "SEQ_ITP",
              c("Type", "Model", "Domain", "InitPrec", "Search", "PredSplit"), "Type",
              log_scale = F, bin_width = 2, pt_position = "jitter")
pairwise_plot(d, "Iterations", "Refinement", "FW_CRAIG_ITP", "SEQ_ITP",
              c("Type", "Model", "Domain", "InitPrec", "Search", "PredSplit"), "Type",
              log_scale = F, bin_width = 2, pt_position = "jitter")
```

## Correlations
```{r}
for(t in unique(d$Type)){
  print(paste("Correlations for type", t))
  d_corr <- d %>% filter(Type == t) %>%
    select(TimeMs, Vars, Size, Iterations, ArgSize, ArgDepth, CexLen)
  tryCatch(
    corrgram(d_corr, order = T, lower.panel = panel.shade, upper.panel = panel.pie),
    error = function(err) {})
}
```


